---
title: 'R Notebook sandbox: Playing with Correlation'
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 4
    fig_caption: true
    number_sections: true
---

```{r,mychunk-common-include,message=FALSE}
library(devtools);
my.source = 'local';
local.path = "/Users/christophermims/_git_/WSU_STATS419_FALL2020/WEEK-06/";
source( paste0(local.path,"functions/libraries.R"), local=T );
```

# Correlation

## Review of Basic Differences Testing 

### Variance
Variance measures the spread of a variable.  Consider the following three examples:

Can't perform operations on WSUHumanVerse because it takes RTools, which is only for windows systems, and I run a mac.

### Iris

```{r, variance-iris, message = FALSE}
iris.df = iris;
iris.df$Species = as.numeric(iris.df$Species);

round( cor(iris.df), digits=2);
```

```{r, variance-iris-plot, message=FALSE}
plot(iris.df);
```

```{r, variance-iris-symnum, message=FALSE}
symnum( cor(iris.df),
       diag = TRUE,
       corr = TRUE,
       cutpoints=c(0.4,0.7,0.9),
       symbols = c(" ",".","*","**")
    );
```

```{r, variance-iris-corrplot, message=FALSE}
corrplot::corrplot( (cor(iris.df)) );
```

```{r, variance-iris-function-of, message=FALSE}
Y = iris.df$Sepal.Length;
X = iris.df$Sepal.Width;
myData = data.frame(cbind(Y,X));

plot(X,Y, xlim=range(X), ylim=range(Y) );
  linear.model = lm(Y~X, myData );
  linear.model.summary = summary(linear.model);
  Y.predicted - predict(linear.model);
par(new=TRUE);
  myMain = paste0("R^2=",
                  round(linear.model.summary$r.squared, digits=4),
                    "; adj.R^2=",
                  round(linear.model.summary$adj.r.squared, digits=4)
                    );
plot(X,Y.predicted, main=myMain, ylab="", xlim=range(X), ylim=range(Y), col="red");
  abline(linear.model, col="blue");
```

```{r, variance-iris-pearson, message=FALSE}
cor(X, Y, method="pearson");
```

```{r, variance-iris-kendall, message=FALSE}
cor(X, Y, method="kendall");
```

```{r, variance-iris-spearman, message=FALSE}
cor(X, Y, method="spearman");
```

```{r, variance-iris-test-pearson, message=FALSE}
cor.test(X, Y, method="pearson");
```

```{r, variance-iris-test-kendall, message=FALSE}
cor.test(X, Y, method="kendall");
```

```{r, variance-iris-test-spearman, message=FALSE}
cor.test(X, Y, method="spearman");
```

```{r, variance-iris-ggplot, message=FALSE}
library(ggplot2);

ggplot() + geom_point(aes(x = X,
                          y = Y), colour = 'red') +
geom_line(aes(x = X,
              y = Y.predicted), colour = 'blue') +
  
ggtitle('X vs Y') +
xlab('X') +
ylab('Y')

```

### Will & Denzel

```{r, will-denzel-setup, message=FALSE}
library(devtools);
path.github = "https://raw.githubusercontent.com/MonteShaffer/humanVerseWSU/master/";
source_url( paste0(path.github, "humanVerseWSU/R/functions-dataframe.R") );

library(devtools);
source_url("http://md5.mshaffer.com/WSU_STATS419/denzel");
  str(denzel);
```

```{r, will-denzel-will, message=FALSE}
source_url("http://md5.mshaffer.com/WSU_STATS419/will");
  str(will);
```

```{r, will-denzel-func, message=FALSE}
imdb.df = rbind(denzel$movies.50, will$movies.50);

imdb.rv = removeAllColumnsBut( imdb.df,
                               c("rank", "year", "minutes", "ratings", "metacritic", "votes", "millions") );

dim(imdb.rv);
```

```{r, will-denzel-play, message=FALSE}
imdb.rv = removeNAsFromDataFrame(imdb.rv, c("metacritic", "millions") );

dim(imdb.rv);
```

```{r, will-denzel-inflation, message=FALSE}
loadInflationData();

imdb.rv = standardizeDollarsInDataFrame(imdb.rv, 2000, "millions", "year", "adjMillions2000");

round( cor(imdb.rv), digits=2);
```

```{r, will-denzel-plot, message=FALSE}
plot(imdb.rv);
```

```{r, will-denzel-symnum, message=FALSE}
symnum( cor(imdb.rv),
        diag = TRUE,
        corr = TRUE,
        cutpoints = c(0.4, 0.7, 0.9), 
        symbols = c(" ", ".", "*", "**")
        );
```

```{r, will-denzel-corrplot, message=FALSE}
corrplot::corrplot( (cor(imdb.rv)) );
```

##### Rank vs. Minutes

This is a negatively correlated pair. It is of medium to low intensity. 

##### Millions vs. Year

This is a positively correlated pair. It is of medium to low intensity.

##### Ratings vs. AdjMillions2000

This is a positively correlated pair. It is a very weak intensity correlation. 

```{r, will-denzel-corr-pair, message=FALSE}
Y = imdb.rv$ratings;
X = imdb.rv$adjMillions2000;

myData = data.frame(cbind(Y, X));

plot(X,Y, xlim=range(X), ylim=range(Y) );
  linear.model = lm(Y~X, myData );
  linear.model.summary = summary(linear.model);
  Y.predicted = predict(linear.model);
par(new=TRUE);
  myMain = paste0("R^2=",
                  round(linear.model.summary$r.squared, digits=4),
                    "; adj.R^2=", 
                  round(linear.model.summary$adj.r.squared, digits=4)
                    );
plot(X, Y.predicted, main=myMain, ylab="", xlim=range(X), ylim=range(Y), col="red");
  abline(linear.model, col="blue");
```

```{r, will-denzel-pearson, message=FALSE}
cor(X,Y, method="pearson");
```

```{r, will-denzel-kendall, message=FALSE}
cor(X, Y, method="kendall");
```

```{r, will-denzel-spearman, message=FALSE}
cor(X, Y, method="spearman");
```

```{r, will-denzel-test-pearson, message=FALSE}
cor.test(X, Y, method="pearson");
```

```{r, will-denzel-test-kendall, message=FALSE}
cor.test(X, Y, method="kendall");
```

```{r, will-denzel-test-spearman, message=FALSE}
cor.test(X, Y, method="spearman");
```

```{r, will-denzel-3Dplot, message=FALSE}
library(rgl);
# Z = imdb.rv$millions2000;
# plot3d(x=X, y=Y, z=Z,
#        type="p", colo="red",
#        xlab="X", ylab="Y", zlab="Z",
#        size=5, lwd=15, box=F
#        );

X = rnorm(100,0,1);
Y = rnorm(100,0,1);
z = X + Y;
plot3d(x=X, y=Y, z=Z,
       type="p", col="red",
       xlab="X", ylab="Y", zlab="Z",
       size=5, lwd=15, box=F
       );

```

#### Measure

```{r, measure-setup, message=FALSE}
library(devtools);
measure.instructor = read.csv("http://md5.mshaffer.com/WSU_STATS419/measure-04343803d489fe8ee2c5f6ab97a101e9.txt", header=TRUE, sep="|");

getOne = c("hand.length", "hand.width", "hand.elbow", "elbow.armpit", "arm.reach", "foot.length", "floor.kneepit", "floor.hip", "floor.armpit");
n.rows = dim(measure.instructor)[1];


for(one in getOne)
  {
  measure.instructor[one] = NA;
  }
  
for(i in 1:n.rows)
  {  
  measure.row = measure.instructor[i,];
  for(one in getOne)
    {
    nidx = getIndexOfDataFrameColumns(measure.instructor, one);
    
    myleft = paste0(one,".left");
      lidx = getIndexOfDataFrameColumns(measure.row, myleft);
    myright = paste0(one,".right");
      ridx = getIndexOfDataFrameColumns(measure.row, myleft);
    
      print(paste0(
                  "left: ",myleft," --> ",lidx,
                  " right: ",myright," --> ",ridx
                  )
            );
      
      row.m = mean(
            c(as.numeric(unlist(measure.row[lidx])),
            as.numeric(unlist(measure.row[ridx]))),
            na.rm=TRUE);
      
    measure.instructor[i,nidx] =  row.m;
    }
  }
```

