---
title: 'Week 03 Assignment 02_datasets'
author: 'Christopher Mims'
date: "'r format(Sys.time(), '%d %B %Y')'"
output: 
  html_document:
    df_print: paged
    toc: true
    toc-float: true
    fig_caption: true
    number_sections: true
my-var: "Christopher"
---


```{r, mychunk-common-include, message = FALSE}
library(devtools); 
my.source = 'local';
local.path = "/Users/christophermims/_git_/WSU_STATS419_FALL2020/";
source( paste0(local.path,"functions/libraries.R"), local=T );

```

# Matrix

Create the "rotate matrix" functions as described in lectures.  Apply to the example "myMatrix".

```{r, mychunk-matrix-include}
local.path = "/Users/christophermims/_git_/WSU_STATS419_FALL2020/";
source( paste0(local.path,"functions/functions-matrix.R"), local=T );

myMatrix = matrix ( c (
                      1, 0, 2,
                      0, 3, 0,
                      4, 0, 5
                      ), nrow = 3, byrow = T);
myMatrix

```

## Transpose Matrix
```{r, mychunk-matrix-transpose}

transposeMatrix = function(matrix) {
  t(matrix);
}

transMatrix = transposeMatrix(myMatrix);
transMatrix
```

## Rotate Matrix
```{r, mychunk-matrix-rotate}

rotateMatrix = function(matrix, clockwise = T) {
  if (clockwise) {
    t(apply(matrix, 2, rev));
  }
  else {
    apply(t(matrix), 2, rev);
  }  
}

rotMatrix = rotateMatrix(myMatrix);
rotMatrix
```

## Rotate Matrix 90 Degrees
```{r, mychunk-matrix-rotate90}

rotateMatrix90= function(matrix, clockwise = T) {
  rotateMatrix(matrix, clockwise);
}

rot90Matrix = rotateMatrix90(myMatrix)
rot90Matrix
```

## Rotate Matrix 180 Degrees
```{r, mychunk-matrix-rotate180}

rotateMatrix180= function(matrix, clockwise = T) {
  rotateMatrix(
    rotateMatrix(
      matrix));
}

rot180Matrix = rotateMatrix180(myMatrix)
rot180Matrix
```

## Rotate Matrix 270 Degrees
```{r, mychunk-matrix-rotate270}

rotateMatrix270= function(matrix, clockwise = T) {
  rotateMatrix(
    rotateMatrix(
      rotateMatrix(
        matrix)));
}

rot270Matrix = rotateMatrix270(myMatrix)
rot270Matrix
```

## Rotate Matrix by Given Angle
```{r, mychunk-matrix-rotate-by-angle}

rotateMatrixAngle= function(matrix, angle = 0, clockwise = T) {
  rem = angle %% 90;
  if (rem != 0) {
    print("Not a valid angle!");
  }
  if (angle < 0) {
    # This will reverse the direction of rotation (from clockwise to counterclockwise)
    clockwise = if(clockwise) FALSE else TRUE;
    angle = abs(angle);
  }
  div = angle %% 360;
  switch(as.character(div),
         "90" = rotateMatrix90(matrix, clockwise),
         "180" = rotateMatrix180(matrix, clockwise),
         "270" = rotateMatrix270(matrix, clockwise),
         matrix
         );
}

rotAngleMatrix = rotateMatrixAngle(myMatrix, 90)
rotAngleMatrix
rotAngleMatrix = rotateMatrixAngle(myMatrix, 180)
rotAngleMatrix
rotAngleMatrix = rotateMatrixAngle(myMatrix, 270)
rotAngleMatrix
rotAngleMatrix = rotateMatrixAngle(myMatrix, 220)
rotAngleMatrix
rotAngleMatrix = rotateMatrixAngle(myMatrix, -90)
rotAngleMatrix
```

#IRIS

Recreate the graphic for the IRIS Data Set using R.  Same titles, same scales, same colors.  See: https://en.wikipedia.org/wiki/Iris_flower_data_set#/media/File:Iris_dataset_scatterplot.svg

```{r, mychunk-iris-demo}

plot(~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width,
	bg=c("red", "green3", "blue")[as.integer(Species)],
	pch=21,
	main="Iris Data (red=setosa,green=versicolor,blue=virginica)",
	data=iris)

```


Sentences: [Write 2-3 sentences concisely defining the IRIS Data Set.  Maybe search KAGGLE for a nice template.  Be certain the final writeup are your own sentences (make certain you modify what you find, make it your own, but also cite where you got your ideas from).  NOTE:  Watch the video, Figure 8 has a +5 EASTER EGG.]

The Iris flower data set is a multivariate data set introduced in 1936 by Ronald Fisher, a statistician and biologist, in his paper "The use of multiple measurements in Taxonomic problems." It consists of four meausrements (sepal length, sepal width, petal length, and petal width) of three species of Iris flower (Iris setosa, Iris versicolor, and Iris virginica). This data set has become the test case for many statistical classification techniques in machine learning.

*Need to cite wikipedia here*

# Personality

Import "personality-raw.txt" into R.  

```{r, mychunk-personality, message = FALSE}
local.path = "/Users/christophermims/_git_/WSU_STATS419_FALL2020/";
source( paste0(local.path,"functions/functions-dataframe.R"), local=T );
df.personality.raw = read.csv("datasets/personality-raw.txt", sep="|", header=T)
```

## Cleanup RAW

Create two new columns from the current column "date_test":  year and week. Stack Overflow may help:  https://stackoverflow.com/questions/22439540/how-to-get-week-numbers-from-dates ... Sort the new data frame by YEAR, WEEK so the newest tests are first ... The newest tests (e.g., 2020 or 2019) are at the top of the data frame.  Then remove duplicates using the unique function based on the column "md5_email".  Save the data frame in the same "pipe-delimited format" ( | is a pipe ) with the headers.  You will keep the new data frame as "personality-clean.txt" for future work (you will not upload it 

```{r, mychunk-personality-ceanup-delete-columns, message = FALSE}
removeColumnsFromDF = function(dataFrame, deleteCol) {
  for(deleteCol in deleteCol) {
    dataFrame[deleteCol] = NULL;
  }
  dataFrame;
}
myCol = "V00";
df.personality.clean = removeColumnsFromDF(df.personality.raw, myCol);
df.personality.clean = separate(df.personality.clean, date_test, c("YEAR", "WEEK"), sep = "\\b\\s\\b", remove = TRUE);
df.personality.clean$WEEK = df.personality.clean$YEAR;
df.personality.clean$YEAR = convertDateStringToFormat(df.personality.clean$YEAR);



```


# Will vs Denzel
```{r, mychunk-imdb-include, message = FALSE}
local.path = "/Users/christophermims/_git_/WSU_STATS419_FALL2020/";
source( paste0(local.path,"functions/functions-imdb.R"), local=T );

```

## Will Smith
```{r, ychunk-will, fig.cap = c("Will Smith scatterplot: IMDB(2020)", "Will Smith boxplot raw millions: IMDB(2020)")}
nmid = "nm0000226";
  will = grabFilmsForPerson(nmid);
  plot(will$movies.50[,c(1,6,7:10)]);
    boxplot(will$movies.50$millions);
    widx = which.max(will$movies.50$millions);
  will$movies.50[widx,];
    summary(will$movies.50$year);

```

## Denzel Washington
```{r,mychunk-denzel}
nmid = "nm0000243";
  denzel = grabFilmsForPerson(nmid);
  plot(denzel$movies.50[,c(1,6,7:10)]);
    boxplot(denzel$movies.50$millions);
    widx = which.max(denzel$movies.50$millions);
  denzel$movies.50[widx,];
    summary(denzel$movies.50$year);

```

## BoxPlot of Top-50 movies using Raw Dollars
```{r,mychunk-boxplot-raw}
par(mfrow=c(1,2));
  boxplot(will$movies.50$millions, main=will$name, ylim=c(0,360), ylab="Raw Millions");
  boxplot(denzel$movies.50$millions, main=denzel$name, ylim=c(0, 360), ylab="RawMillions");

  par(mfrow=c(1,1));

```

## Side-by-Side Comparisons

### Adjusted Dollars (2000)
### Total Votes (Divide by x,xxx)
### Average Ratings